---
title: "MATH2349 Semester 2, 2018"
author: "Student name(s) and numbers comes here"
subtitle: Assignment 2
output:
  html_notebook: default
---


## IMPORTANT NOTE: 

All the instructions to complete this assignment are available on the [MATH2349_1850 Assignment_2](https://drive.google.com/open?id=1kBnzsf6I52moGGVlQwPE0HQBNoekemwWZlQNTN8W8pw) Word file. Please read through this document carefully before submitting your report.

**Groups**

Students are permitted to work individually or in groups of up to 3 people for Assignment 2. **Each group must fill out the [group registration form](https://docs.google.com/forms/d/e/1FAIpQLSfqdJEW7vK7odCCWDX1FEzZZKwmgDyFieWmVv7whGfil-OoiA/viewform?usp=sf_link)** before 9/09/2018 to register their group details. 

This report must be uploaded to Turnitin as a PDF with your code chunks and outputs showing. The easiest way to achieve this is to Preview your notebook in HTML (by clicking Preview) → Open in Browser (Chrome) → Right click on the report in Chrome → Click Print and Select the Destination Option to Save as PDF. 


**DELETE the instructional text provided in the template. Failure to do this will INCREASE the SIMILARITY INDEX reported in TURNITIN**. If you have any questions regarding the assignment instructions and the R template, please post it on [Slack](https://math2349-1850.slack.com) under the `#assignment2` channel. 


## Setup

Install and load the necessary packages to reproduce the report here:

```{r, echo = TRUE, message=FALSE}

# This is a chunk where you can load the necessary packages required to reproduce the report. Here are some example packages, you may add others if you require  
install.packages("outliers")
library(readr)
library(tidyr)
library(dplyr)
library(Hmisc)
library(outliers)

```


## Read WHO Data

Read the WHO data using an appropriate function.

```{r echo=TRUE}

# This is an R chunk for reading the WHO data. Provide your R codes here:

setwd("C:\\Users\\marga\\OneDrive\\Documents\\Uni\\Data Preprocessing\\Assignment 2")
who <- read_csv("WHO.csv")

head(who)

```




## Tidy Task 1:

Use appropriate `tidyr` functions to reshape the WHO data set into the form given in the MATH2349_1810 Assignment_2 instructions document. 


```{r, echo=TRUE}

# This is an R chunk for tidy task 1. Provide your R codes here:
who_tidy1 <- who %>% gather(code, value, new_sp_m014:new_rel_f65)

head(who_tidy1)

```



## Tidy Task 2:

The WHO data set is not in a tidy format yet. The `code` column still contains four different variables' information (see variable description section for the details). Separate the `code` column and form four new variables using appropriate `tidyr` functions.  The final format of the WHO data set for this task should be in the form given in the MATH2349_1810 Assignment_2 instructions document.  


```{r echo=TRUE}

# This is an R chunk for tidy task 2. Provide your R codes here:


who_tidy2 <- who_tidy1 %>% separate(code, into = c("new", "var", "sex_age"), sep = "_") %>%
                            mutate(sex = substr(sex_age, 0,1), age = substr(sex_age, 2, 4 )) %>%
                            select(country, iso2, iso3, year, new, var, sex, age, value) 
                            
                              
                             
head(who_tidy2)


```


## Tidy Task 3:

The WHO data set is not in a tidy format yet. The `rel`, `ep`, `sn`, and `sp` keys need to be in their own columns as we will treat each of these as a separate variable. In this step, move the `rel`, `ep`, `sn`, and `sp` keys into their own columns. The final version of the WHO data set for this task should be in the form given in the MATH2349_1810 Assignment_2 instructions document.  


```{r echo=TRUE}

# This is an R chunk for tidy task 3. Provide your R codes here:

who_tidy3 <- who_tidy2 %>%  spread(key = var, value = value)

head(who_tidy3)

```


## Tidy Task 4: 

There is one more step to tidy WHO data set. We have two categorical variables `sex` and `age`. Use `mutate()` to factorise sex and age. For age variable, you need to create labels and also order the variable. Labels would be: <15, 15-24, 25-34, 35-44, 45-54, 55-64, 65>=. The final tidy version of the WHO data set would be in the form given in the MATH2349_1810 Assignment_2 instructions document.


```{r echo=TRUE}

# This is a chunk for Task 4. Provide your R codes here: 

who_tidy4 <- who_tidy3 %>% mutate(sex = factor(sex, 
                                               levels = c("m", "f"),
                                               labels = c("m", "f")
                                               ),
                                  age = factor(age, 
                                               levels = c("014", "1524", "2534", "3544", "4554", "5564", "65"),
                                               labels = c("<15", "15-24", "25-34", "35-44", "45-54", "55-64", "65>="),
                                               ordered = TRUE))


head(who_tidy4)

```

## Task 5: Filter & Select


Drop the redundant columns `iso2` and `new`, and filter any three countries from the tidy version of the WHO data set. Name this subset of the data frame as `WHO_subset`.

```{r}

# This is a chunk for Task 5. Provide your R codes here: 

WHO_subset <- who_tidy4 %>% select(-(iso2), -(new)) %>%
                            filter(country == "Fiji" | country == "Samoa" | country == "Tonga")

WHO_subset


```



## Read Species and Surveys data sets

Read the Species and Surveys data sets using an appropriate function. Name these data frames as `species` and `surveys`, respectively.

```{r}

# This is an R chunk for reading the Species and Surveys data sets. Provide your R codes here:
species <- read_csv("species.csv")
surveys <- read_csv("surveys.csv")

head(species)
head(surveys)

```




## Task 6: Join  

Combine `surveys` and `species` data frames using the key variable `species_id`. **For this task, you need to add the species information (`genus`, `species`, `taxa`) to the `surveys` data**.  Rename the combined data frame as `surveys_combined`.


```{r}

# This is a chunk for Task 6. Provide your R codes here: 

surveys_combined <- surveys %>% left_join(species, by = "species_id")

surveys_combined

```



## Task 7: Calculate 

Using the `surveys_combined` data frame, calculate the average weight and hindfoot length of one of the species observed in each month (irrespective of the year). Make sure to exclude missing values while calculating the average.


```{r}

# This is a chunk for Task 7. Provide your R codes here: 

megalotis_avg <- surveys_combined %>% filter(species == "megalotis") %>%
                    group_by(month) %>%
                    summarise(mean_weight = mean(weight, na.rm = TRUE),
                              mean_foot = mean(hindfoot_length, na.rm = TRUE)
                                )
                    

megalotis_avg


```



## Task 8: Missing Values

Select one of the years in the `surveys_combined` data frame, rename this data set as `surveys_combined_year`. Using `surveys_combined_year` data frame, find the total missing values in `weight` column grouped by `species`. Replace the missing values in `weight` column with the mean values of each species. Save this imputed data as `surveys_weight_imputed`.


```{r echo=TRUE}

# This is a chunk for Task 8. Provide your R codes here: 

surveys_combined_year <- surveys_combined %>% filter(year == "1988") 

surveys_combined_year %>% group_by(surveys_combined_year$species) %>% summarise(na_count = sum(is.na(weight)))  

surveys_weight_imputed <- surveys_combined_year %>% group_by(surveys_combined_year$species) %>% mutate(weight = ifelse(is.na(weight), mean(weight, na.rm = TRUE), weight))

surveys_weight_imputed 
```


## Task 9: Inconsistencies or Special Values

Inspect the `weight` column in `surveys_weight_imputed` data frame for any further inconsistencies or special values (i.e., NaN, Inf, -Inf) . Trace back and explain briefly why you got such a value.

```{r}

# This is a chunk for Task 9. Provide your R codes here: 

is.nan(surveys_weight_imputed$weight)

is.special <- function(x){
if (is.numeric(x)) !is.finite(x) else is.na(x)
}


is.special <- function(x){
if (is.numeric(x)) !is.finite(x)
}


sapply(surveys_weight_imputed$weight, is.special)



```

Explanation: Some species in the data frame have no values available for weight for any observation of that species. When calculating the mean by species, we have excluded NAs so any species with no weight values recorded will have no valid numbers available for calculation. Therefore, the final output includes NaN (not a number) results for the weights of species with no valid weight observation. 


## Task 10: Outliers

Using the `surveys_combined` data frame, inspect the variable hindfoot length for possible univariate outliers. If you detect any outliers use any of the methods outlined in the Module 6 notes to deal with them. Explain briefly the actions that you take to handle outliers. 

```{r}

# This is a chunk for Task 10. Provide your R codes here: 

install.packages("DescTools")
library(DescTools)

surveys_combined$hindfoot_length %>% boxplot(main = "Hindfoot Length", ylab = "cm", col = "grey")

summary(surveys_combined$hindfoot_length)



```
Explanation: Capping outliers replaces the outliers with the nearest values that are not outliers. As there are only two outliers in a set of 35,549 obvservations, it is highly possible that these extreme observations are the result of error and not truely reflective of the general population. 


<br>
<br>
