---
title: "MATH2349 Semester 2, 2018"
author: "Meg Cuddihy (s3608125), Sam Holt (s3381728), Verity Miles (s3644459)"
subtitle: Assignment 2
output:
  html_notebook: default
---


### Setup

Install and load the necessary packages to reproduce the report here:

```{r, echo = TRUE, message=FALSE}

# Loading all the required packages

library(readr)
library(tidyr)
library(dplyr)
library(Hmisc)
library(outliers)
library(DescTools)
library(kableExtra)
library(mosaic)

```


### WHO data

Importing the WHO data

```{r, message = FALSE}

who <- read_csv("Who.csv")
kable(head(who[,1:5]), caption = "WHO data preview") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: all code the same


### Tidy Task 1:

Use tidyr functions to reshape/ gather the WHO data

```{r, echo=TRUE}

who_tidy1 <- who %>% gather(code, value, 5:length(who))
kable(head(who_tidy1)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: Sam's code


### Tidy Task 2:

Splitting/ separating the code column to get closer to tidy data

```{r}

who_tidy2 <- who_tidy1 %>% 
  separate(code, c("new", "var", "sex"), sep = "_") %>% 
  separate(sex, c("sex", "age"), sep = 1)

kable(head(who_tidy2)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: Verity's code


### Tidy Task 3:

Spreading the data to get it into a tidy format

```{r}

who_tidy3 <- who_tidy2 %>% spread(var, value)

kable(head(who_tidy3)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: all code the same


### Tidy Task 4: 

Mutating and factorising variables

```{r}

who_tidy4 <- who_tidy3 %>% mutate(sex = factor(sex, levels = c("m", "f"), labels = c("m", "f")),
                                  age = factor(age, levels = c("014", "1524", "2534", "3544", "4554", "5564", "65"),
                                               labels = c("<15", "15-24", "25-34", "35-44", "45-54", "55-64", "65>="),
                                               ordered = TRUE))

#Checking the class of sex and age
sapply(who_tidy4[c("sex", "age")], class)

kable(head(who_tidy4)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: Meg's code


### Task 5: Filter & Select

Filtering and selecting the WHO data

```{r}

countries <- c("Fiji", "Germany", "Korea")
who_subset <- who_tidy4 %>% select(-c(2, 5)) %>% 
  filter(country %in% countries)

kable(head(who_subset)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: Verity's code


## Species and Surveys data sets

Importing the species and surveys data

```{r, message = FALSE}

surveys <- read_csv("surveys.csv")
species <- read_csv("species.csv")

kable(head(surveys), caption = "Surveys data preview") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

kable(head(species), caption = "Species data preview") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: all code the same


### Task 6: Join  

Join the species and surveys data

```{r}

surveys_combined <- surveys %>% left_join(species[, c("species_id", "genus", "species", "taxa")], by = "species_id")

kable(head(surveys_combined)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: Verity's code


### Task 7: Calculate 

Calculating mean weight and hindfoot lengths

```{r}

species_avg <- surveys_combined %>% filter(species == "fulviventer") %>% 
  group_by(month) %>% 
  summarise(mean_weights = mean(weight, na.rm = TRUE),
            mean_foot = mean(hindfoot_length, na.rm = TRUE))

kable(head(species_avg)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

```
Note: mainly Meg's code


### Task 8: Missing Values

Dealing with missing values through imputing

```{r}

surveys_combined_year <- surveys_combined %>% filter(year == "1990") 
surveys_combined_year %>% group_by(species) %>% summarise(Missing = sum(is.na(weight)))
surveys_weight_imputed <- surveys_combined_year %>%  group_by(species) %>% mutate(weight = ifelse(is.na(weight), mean(weight, na.rm = TRUE), weight))

```
Note: Sam's code

There are still missing values as there are some entire species that only have NA values for weight in 1990. The means of these species is NaN.


### Task 9: Inconsistencies or Special Values

Checking for inconsistencies in the weight data and explaining why they have occurred.

```{r}

sum(is.nan(surveys_weight_imputed$weight))

is.special <- function(x){
if (is.numeric(x)) !is.finite(x) else is.na(x)
}

any(sapply(surveys_weight_imputed$weight, is.special))

#Checking one example
example <- first(which(is.na(surveys_weight_imputed$weight)))
species_eg <- surveys_weight_imputed[example,]$species
no_data_species <- filter(surveys_weight_imputed, species == species_eg)
favstats(no_data_species$weight) #Shows that all values for this species are NaN

surveys_weight_imputed %>% group_by(species) %>% summarise(Missing = sum(is.na(weight)), Mean = mean(weight))

```
Note: combination of code

Explanation: Some species in the data frame have no values available for weight for any observation of that species. When calculating the mean by species, we have excluded NAs so any species with no weight values recorded will have no valid numbers available for calculation. Therefore, the final output includes NaN (not a number) results for the weights of species with no valid weight observation. 


### Task 10: Outliers

Checking for outliers in the hindfoot length data

```{r}

hist(surveys_combined$hindfoot_length, breaks = 30,
     main = "Histogram of Hind Foot Length",
     xlab = "Hind Foot Length",
     col = "#39b3a7") #does not seem to be normally distributed
x <- boxplot(surveys_combined$hindfoot_length,
             main = "Hind Foot Length") #hence the Tukey method is used
x$out
IQR(surveys_combined$hindfoot_length, na.rm = TRUE) * 3
# Winsorising the outliers as they don't follow a normal distribution nor are they extreme (neither exceed 3*IQR)

# Winsorising with the Winsorize() function from the DescTools Package
surveys_combined_capped <- Winsorize(surveys_combined$hindfoot_length, na.rm = TRUE)
summary(surveys_combined_capped)
boxplot(surveys_combined_capped,
        main = 'Hindfoot Length After Winsorising',
        ylab = 'Hindfoot Length')

hist(surveys_combined_capped, breaks = 30,
     main = "Histogram of Capped Hind Foot Length",
     xlab = "Capped Hind Foot Length",
     col = "#39b3a7") #show that there are now no outliers but more values at each edge of the data

```
Note: Sam's code

The hindfoot length data is not normally distributed as shown in the initial histogram. The boxplot indicates that there are two outliers present. The Tukey method was used to transform the data. Rather than ignoring the outliers or setting them to the mean value, we chose to cap/ winsorise them. We did this as they don't follow a normal distribution nor are they very extreme. The second histogram shows that the outliers were dealt with as the spread of the data is not as wide and there are more values in the first and last bins than in the original histogram.
